sex

### 2021-08-27 Manuscript/Manuscript.tex
cx

A Palmer amaranth accession (Per1) from Perkins County, Nebraska with no
reported herbicide resistance according to Oliveira et al. (2021) was
selected for this study. Three weeks prior to the establishment of each
cohort, seeds were planted in plastic trays containing potting-mix.
Emerged seedlings (1 cm) were transplanted into 200
cm\textsuperscript{-3} plastic pots (a plant pot\textsuperscript{-1}).
Palmer amaranth seedlings were supplied with adequate water and kept
under greenhouse conditions at the University of Wisconsin-Madison,
University of Nebraska-Lincoln, and Western Illinois University; and
kept outdoors at the Perkins extension office in Grant, NE until the 2-3
leaf stage (5 to 8 cm height) when they were transported to the field.

@@

\caption{\label{tab:unnamed-chunk-2}Field study attributes }
\centering
\resizebox{\linewidth}{!}{
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lllllll}
\toprule
Attributes &  & Arlington, WI & Clay Center, NE & Grant, NE & Lincoln, NE & Macomb, IL\\
\midrule
Bareground &  &  &  &  &  & \\

 & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\

@@

 & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\

 & Stage at 1 cohort & V2-3 &  &  &  & \\

 & Stage at 2 cohort & V6-7 &  &  &  & \\

@@

\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\

@@

 & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\

 & Stage at 1 cohort & V1-2 &  &  &  & \\

 & Stage at 2 cohort & V5-6 &  &  &  & \\

@@

\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
\end{tabular}}
\end{table}

The field experimental units were three adjacent 9.1 m wide (12 rows at
76.2 cm row spacing) by 10.7 m long. The experimental design were
arranged in factorial design with three crops, two transplanting times
simulating two cohorts, repeated across five locations. Each filed
experimental unit was planted with corn, soybean, or kept under
bareground. The two transplanting timings were June 1 (first cohort) and
July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight
plants were equidistantly placed (0.76 m apart) between rows within each
crop. After a week, one plant was eliminated and one was kept, resulting
in 24 plants per experimental unit and transplanting time (Figure 2).
When needed, Palmer amaranth plants were supplied with water during the
first week after transplanting to assure seedling survival.

@@

Analyses of Palmer amaranth height and biomass were performed with a
linear mixed model using \emph{lmer} function from ``lme4'' package
(Bates et al., 2015). Plant height and biomass were log transformed to
meet model assumption of normality. In the model, crop (bareground,
corn, soybean) and cohort time (first and second) were the fixed effects
and year nested with location the random effects. Analysis of variance
was performed with \emph{anova} function from ``car'' package (Fox and
Weisberg, 2018). Marginal means and compact letter display were
estimated with \emph{emmeans} and \emph{cld} from packages ``emmeans''
(Lenth et al., 2021) and ``multcomp'' (Hothorn et al., 2008),
respectively.

@@

A binary logistic regression was fitted to Palmer amaranth sex
dimorphism. Binary logistic regression is used for predicting binary
classes (Bangdiwala, 2018), such as the probability of a plant being
female in a dioecious species. Prior to the analysis, missing values
were removed from the dataset. Also, data from Grant was not used in
this analysis due to the uniform plant harvesting at that location. The
complete dataset was splitted into 80\% train and 20\% test data. The
80\% train is used for the model training and the 20\% test is used for
checking model performance on unseen dataset. With 80\% dataset, a
generalized linear model (base R \emph{glm} function) was fitted to
binary response variable, the probability of being female (0 to male and
1 to female). The independent variables were day of year harvest,
height, weight, and crop (without interaction). The model family was
binomial with a logit function. The model fit was assessed through
pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and
likelihood ratio using \emph{nagelkerke} function from ``rcompanion''
package (Mangiafico, 2021). The marginal effects computation was
performed with Average Marginal Effects (AMEs) at every observed value
of x and average across the results (Leeper, 2017) using \emph{margins}
function from ``margins'' package (Leeper et al., 2021). The 20\% test
data was predicted using the \emph{predict} function with a cutoff
estimation for male or female using \emph{performance} function. The
model quality prediction from the classification algorithm was measured
with precision (\emph{precision} function), recall (\emph{recall}
function) and F1-score (\emph{f\_meas} function) using the ``yardstick''
package (Kuhn et al., 2021). The precision determines the accuracy of
positive predictions (female plants), recall determines the fraction of
positives that were correctly identified, and F1-score is a weighted
harmonic mean of precision and recall with the best score of 1 and the
worst score of 0. F1-score conveys the balance between the precision and
the recall (Yacouby and Axman, 2020). The area under the curve was also
estimated with XXXXXXXXXXXXXXXXX.

@@

\caption{\label{tab:unnamed-chunk-3}Average marginal effects of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrrr}
\toprule
Term & AME & SE & Lower & Upper & Z-score & P-value\\
\midrule
crop\_bareground & -0.048 & 0.054 & -0.154 & 0.059 & -0.876 & 0.381\\

@@

Palmer amaranth grow and development in second cohort was limited due to
the crop competitive ability at advanced development stages. Plants were
transplanted at greater crop height, which reduced Palmer amaranth
competitiveness. As a result, Palmer amaranth height and biomass was
lower compared to its first cohort. Palmer amaranth growing without crop
competition produced the highest amounts of biomass. The Palmer amaranth
strategy in bareground was to invest biomass in growing plant width and
height. Nonetheless, Palmer amaranth produced 17\% less biomass in
second cohort compared to first cohort timing. In a bareground study,
early emerged Palmer amaranth without competition was 50\% taller than
late emerged plants (Webster and Grey, 2015). These results suggest that
crop competition is not the only factor limiting late Palmer amaranth
establishment. The limited growth of Palmer amaranth at second cohort is
likely a reduced plant response to day length, light availability and
thermal units (e.g, growing degree days). The \emph{Amaranthus} species
are sensitive to photoperiod (Wu and Owen, 2014). It is hypothesize that
reduced day length contributed to smaller plants at second cohort
regardless the crop. A study in North Carolina and Illinois predicted
that less than 10\% Palmer amaranth seedlings emergence occurred after
June (Piskackova et al., 2021). In addition, Palmer amaranth negative
impact on soybean (Korres et al., 2020) and cotton (Webster and Grey,
2015) yields was higher when plants were established near to crop
planting.

@@

An ecological approach to reduce Palmer amaranth seed production is
understanding plant biology, including flowering pattern. Our study
suggests that Palmer amaranth flowering was slightly influenced by crops
and cohort timings. Palmer amaranth growing in bareground and corn
resulted in the overall shortest and longest flowering window,
respectively. The shortest flowering window of second cohorts of Palmer
amaranth growing in bareground highlighted the impact of photoperiod on
flowering. When growing in soybean, Palmer amaranth flowering window was
similar to bareground at first cohort but similar to corn at second
cohort timing. Plant flowering initiation is complex and depends on
biological and ecological factors (Lang, 1965). We hypothesize that when
growing in high competition (e.g., second cohort), Palmer amaranth
plants tend to initiate flowering early, as well as having an extended
flowering window. A study has shown that non-native Palmer amaranth
initiated flowering two weeks prior to the native waterhemp in Iowa
(Baker, 2021). Early flower initiation could also be a plant strategy
when growing in stress conditions. For example, when growing under water
stress, early flowering in Palmer amaranth resulted in a mismatch
between female and male plants by seven days (Mesgaran et al., 2021). A
mismatch in Palmer amaranth male and female flowering period can
minimize plant outcross, reducing plant seed production and/or exchange
of resistant alleles (Jhala et al., 2021). Sex dimorphism manipulation
is considered a potential ecological pest control (McFarlane et al.,
2018; Schliekelman et al., 2005).

The mechanisms of sex-determination in plant species is intriguing and
arouse the curiosity of many scientists, including Charles Darwin
(Darwin, 1888). In our study, the sex dimorphism model performance was
decent (AIC 0.64) considering the biology of plant flowering. A 1:1 male
and female sex ratio is a general evolutionary stable strategy for plant
species perpetuation (Fisher, 1930). However, a slight deviation from
1:1 sex ratio might occur in some dioecious species. For example, the
dioecious \emph{Halophila stipulacea} is a female-biased plant in its
native habitat, but the naturalized \emph{H. stipulacea} have a 1:1
ratio (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
female-male ratio to expand into its non-native habitat (Nguyen et al.,
2018). Also, biotic and/or abiotic stress can influence plant sex
determination. Palmer amaranth male-to-female ratio was greater under
high plant densities (Korres and Norsworthy, 2017) and after herbicide
application (Rumpa et al., 2019). Our model estimated that late
flowering, heavier and taller Palmer amaranth plants deviated from 1:1
ratio in favor to female plants. It was reported that female Palmer
amaranth plants invested more in height, stem and biomass while male
invested more in leaf area and leaf dry weight under nutrient deficiency
(Korres et al., 2017). Our model also estimated more female plants in
soybean and bareground compared to corn, which might be linked to plant
competition strategy in each crop. Our results showed the influence of
life-history and ecological traits on sexual dimorphism in Palmer
amaranth. Sexual dimorphism is documented in other dioecious species
(Barrett and Hough, 2013). For example, stronger female plant
competition and greater male tolerance to herbivory was reported in
\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019).
Research on candidate genes for sex determination in \emph{Amaranthus}
species are currently underway but it is far from complete (Montgomery
et al., 2021, 2019). Further studies are also needed to understand the
ecological basis of Palmer amaranth flowering, including the plant
behavior under climate change.

Our study demonstrated the Palmer amaranth plasticity to grow and
develop into arable land of US Midwest. Is likely that Palmer amaranth
range will continue to expanding into new geographies. The migration of
Palmer amaranth into the US Midwest will reshape the landscape as
waterhemp and Palmer amaranth will share the same habitat. The presence
of Palmer amaranth and waterhemp will increase the weed management
complexity. Therefore, preventive management is a priority to minimizing
Palmer amaranth dispersal. Reactive management should focus on
early-season management programs, which would have a large negative
effect on Palmer amaranth growth and development. Long-term tactics that
promote early-season crop advantage against Palmer amaranth, including
diversity of crops in rotation, early/late crop planting, plant width,
and crop residue (e.g.~cover crops) would minimize the negative impact
of Palmer amaranth to cropping systems. The aggressiveness and
differential Palmer amaranth adaptation to agroecosystem is striking and
require national efforts to minimize the species impact on
sustainability and profitability of cropping systems.

@@

@@

@@

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}

\}

\textbackslash caption\{The probability (P) of being female Palmer
amaranth by day of year (A), weight (B), and height (C). Back line
represents the model estimation and shade green the 95\% confidence
intervals\}\label{fig:Figure-5} \textbackslash end\{figure\}

>>>>

A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no
glyphosate resistance according to Oliveira et al. (2021) was selected
for this study. Three weeks prior to the establishment of each cohort,
seeds were planted in plastic trays containing potting-mix. Emerged
seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3}
plastic pots (a plant pot\textsuperscript{-1}). Palmer amaranth
seedlings were supplied with adequate water and kept under greenhouse
conditions at the University of Wisconsin-Madison, University of
Nebraska-Lincoln, and Western Illinois University; and kept outdoors at
the Perkins extension office in Grant, NE until the 2-3 leaf stage (5 to
8 cm height) when they were transported to the field.

@@

\caption{\label{tab:unnamed-chunk-2}Field study attributes from Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}
\centering
\resizebox{\linewidth}{!}{
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lllllll}
\toprule
Attributes &  & Arlington, WI & Clay Center, NE & Grant, NE & Lincoln, NE & Macomb, IL\\
\midrule
Bareground &  &  &  &  &  & \\

 & Weed control & glyphosate\textsuperscript{a} / \emph{S}-metolachor\textsuperscript{b} & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{c} &  &  & \\

@@

 & Weed control & glyphosate / \emph{S}-metolachor & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{d} &  &  & \\

 & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\

 & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\

@@

\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  &  & \\

@@

 & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\

 & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\

 & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\

@@

\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} \emph{S}--metolachor, 1324 g ai ha; \textsuperscript{c} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha; \textsuperscript{d} \emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha}\\
\end{tabular}}
\end{table}

The field experimental units were three adjacent 9.1 m wide (12 rows at
76.2 cm row spacing) by 10.7 m long. The experimental design were
arranged in factorial design with three crops, two transplanting times
simulating two cohorts, repeated across five locations. Each filed
experimental unit was planted with corn, soybean, or kept under
bareground. The two transplant timings were June 1 (first cohort) and
July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight
plants were equidistantly placed (0.76 m apart) between rows within each
crop. After a week, one plant was eliminated and one was kept, resulting
in 24 plants per experimental unit and transplanting time (Figure 2).
When needed, Palmer amaranth plants were supplied with water during the
first week after transplanting to assure seedling survival.

@@

Analyses of Palmer amaranth height and biomass were performed with a
linear mixed model using \emph{lmer} function from ``lme4'' package
(Bates et al., 2015). Plant height and biomass were log transformed to
meet model assumption of normality. In the model, crop (bareground,
corn, soybean) and cohort time (first and second) were the fixed effects
and year nested with location the random effects. Analysis of variance
at \(\alpha\) 0.05 was performed with \emph{anova} function from ``car''
package (Fox and Weisberg, 2018). Marginal means and compact letter
display were estimated with \emph{emmeans} and \emph{cld} from packages
``emmeans'' (Lenth et al., 2021) and ``multcomp'' (Hothorn et al.,
2008), respectively.

@@

A binary logistic regression was fitted to Palmer amaranth sex
dimorphism. Binary logistic regression is used for predicting binary
classes (Bangdiwala, 2018), such as the probability of a plant being
female in a dioecious species. Prior to the analysis, missing values
were removed from the dataset. Also, data from Grant was not used in
this analysis due to the uniform plant harvesting at that location. The
complete dataset was splitted into 80\% train and 20\% test data. The
80\% train is used for the model training and the 20\% test is used for
checking model performance on unseen dataset. With 80\% dataset, a
generalized linear model (base R \emph{glm} function) was fitted to
binary response variable, the probability of being female (0 to male and
1 to female). The independent variables were day of year harvest,
height, weight, and crop (without interaction). The model family was
binomial with a logit function. The model fit was assessed through
pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and
likelihood ratio using \emph{nagelkerke} function from ``rcompanion''
package (Mangiafico, 2021). The marginal effects computation was
performed with Average Marginal Effects (AMEs) at every observed value
of x and average across the results (Leeper, 2017) using \emph{margins}
function from ``margins'' package (Leeper et al., 2021). The 20\% test
data was predicted using the \emph{predict} function with a cutoff
estimation for male or female using \emph{performance} function from
ROCR package (Sing et al., 2005). The model quality prediction from the
classification algorithm was measured with precision (\emph{precision}
function), recall (\emph{recall} function) and F1-score (\emph{f\_meas}
function) using the ``yardstick'' package (Kuhn et al., 2021). The
precision determines the accuracy of positive predictions (female
plants), recall determines the fraction of positives that were correctly
identified, and F1-score is a weighted harmonic mean of precision and
recall with the best score of 1 and the worst score of 0 (Raoniar,
2021). F1-score conveys the balance between the precision and the recall
(Yacouby and Axman, 2020). The area under the receiver operating curve
(AUC-ROC) was also estimated with performance function using the true
positive and false positive rates. Higher the AUC, better the model is
at distinguishing between female and male Palmer amaranth.

@@

\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrrr}
\toprule
Term & AME\textsuperscript{a} & SE\textsuperscript{b} & Lower & Upper & Z-score & P-value\\
\midrule
crop\_bareground & -0.048 & 0.054 & -0.154 & 0.059 & -0.876 & 0.381\\

@@

Palmer amaranth grow and development in second cohort was limited due to
the crop competitive ability at advanced development stages. Plants were
transplanted at greater crop height and width, which reduced Palmer
amaranth competitiveness. As a result, Palmer amaranth height and
biomass was lower compared to its first cohort. Moreover, Palmer
amaranth growing without crop competition produced the highest amounts
of biomass. The Palmer amaranth strategy in bareground was to invest
biomass in growing plant width and height. Nonetheless, Palmer amaranth
produced 17\% less biomass in second cohort compared to first cohort
timing. In a bareground study, early emerged Palmer amaranth without
competition was 50\% taller than late emerged plants (Webster and Grey,
2015). These results suggest that crop competition is not the only
factor limiting late Palmer amaranth establishment. The limited growth
of Palmer amaranth at second cohort is likely a reduced plant response
to day length, light availability and thermal units (e.g, growing degree
days). The \emph{Amaranthus} species are sensitive to photoperiod (Wu
and Owen, 2014). It is hypothesize that reduced day length contributed
to smaller plants at second cohort regardless the crop. A study in North
Carolina and Illinois predicted that less than 10\% Palmer amaranth
seedlings emergence occurred after June (Piskackova et al., 2021). In
addition, Palmer amaranth negative impact on soybean (Korres et al.,
2020) and cotton (Webster and Grey, 2015) yields was higher when plants
were established near to crop planting. Therefore, early management is a
key strategy to minimize the damaging impact of Palmer amaranth to US
Midwest cropping systems.

@@

An ecological approach to reduce Palmer amaranth seed production is
understanding plant biology, including flowering pattern. Our study
suggests that Palmer amaranth flowering was slightly influenced by crops
and cohort timings. Palmer amaranth growing in bareground and corn
resulted in the overall shortest and longest flowering window,
respectively. The shortest flowering window of second cohorts of Palmer
amaranth growing in bareground highlighted the impact of photoperiod on
flowering. When growing in soybean, Palmer amaranth flowering window was
similar to bareground at first cohort but similar to corn at second
cohort timing. Plant flowering initiation is complex and depends on
biological and ecological factors (Lang, 1965). We hypothesize that when
growing in high competition (e.g., second cohort), Palmer amaranth
plants tend to initiate flowering early, as well as having an extended
flowering window. A study has shown that Palmer amaranth initiated
flowering two weeks prior to the native waterhemp in Iowa (Baker, 2021).
Early flower initiation could also be a plant strategy when growing in
stress conditions. For example, when growing under water stress, early
flowering in Palmer amaranth resulted in a mismatch between female and
male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
amaranth male and female flowering period can minimize plant outcross,
reducing plant seed production and/or exchange of resistant alleles
(Jhala et al., 2021). Sex dimorphism manipulation is considered a
potential ecological pest control (McFarlane et al., 2018; Schliekelman
et al., 2005).

The mechanisms of sex-determination in plant species is intriguing and
arouse the curiosity of many scientists, including Charles Darwin
(Darwin, 1888). In our study, the sex dimorphism model performance was
decent (AIC 0.64) considering the biology of plant flowering. A 1:1 male
and female sex ratio is a general evolutionary stable strategy for plant
species perpetuation (Fisher, 1930). However, a slight deviation from
1:1 sex ratio might occur in some dioecious species. For example, the
dioecious \emph{Halophila stipulacea} is a female-biased plant in its
native habitat, but the naturalized \emph{H. stipulacea} have a 1:1
ratio (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
female-male ratio to expand into its non-native habitat (Nguyen et al.,
2018). Also, biotic and/or abiotic stress can influence plant sex
determination. Palmer amaranth male-to-female ratio was greater under
high plant densities (Korres and Norsworthy, 2017) and after herbicide
application (Rumpa et al., 2019). Our model estimated that late
flowering, heavier and taller Palmer amaranth plants slighly deviated
from 1:1 ratio in favor to female plants. It was reported that female
Palmer amaranth plants invested more in height, stem and biomass while
male invested more in leaf area and leaf dry weight under nutrient
deficiency (Korres et al., 2017). Our model also estimated more female
plants in soybean and bareground compared to corn, which might be linked
to plant competition strategy in each crop. Our results showed the
influence of life-history and ecological traits on sexual dimorphism in
Palmer amaranth. Sexual dimorphism is documented in other dioecious
species (Barrett and Hough, 2013). For example, stronger female plant
competition and greater male tolerance to herbivory was reported in
\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019).
Research on candidate genes for sex determination in \emph{Amaranthus}
species are currently underway but it is far from complete (Montgomery
et al., 2021, 2019). Further studies are also needed to understand the
ecological basis of Palmer amaranth flowering, including the plant
behavior under climate change.

Our study demonstrated the Palmer amaranth plasticity to grow and
develop into arable land of US Midwest. It is likely that Palmer
amaranth range will continue to expanding into new geographies. The
migration of Palmer amaranth into the US Midwest will reshape the
landscape as waterhemp and Palmer amaranth will share the same habitat.
The presence of Palmer amaranth and waterhemp will increase the weed
management complexity. Therefore, preventive management is a priority to
minimizing Palmer amaranth dispersal. Reactive management should focus
on early-season management programs, which would have a large negative
effect on Palmer amaranth growth and development. Long-term tactics that
promote early-season crop advantage against Palmer amaranth, including
diversity of crops in rotation, early/late crop planting, plant width,
and crop residue (e.g.~cover crops) would minimize the negative impact
of Palmer amaranth to cropping systems. The aggressiveness and
differential Palmer amaranth adaptation to agroecosystem is striking and
require national efforts to minimize the species impact on
sustainability and profitability of cropping systems.

@@

\leavevmode\hypertarget{ref-raoniar2021}{}%
Raoniar, R. (2021). Modelling {Binary Logistic Regression Using R}
(research-oriented modelling and interpretation). Available at:
\url{https://bit.ly/3BhucN3} {[}Accessed August 11, 2021{]}.

@@

\leavevmode\hypertarget{ref-sing2005}{}%
Sing, T., Sander, O., Beerenwinkel, N., and Lengauer, T. (2005). {ROCR}:
Visualizing classifier performance in {R}. \emph{Bioinformatics} 21,
3940--3941.
doi:\href{https://doi.org/10.1093/bioinformatics/bti623}{10.1093/bioinformatics/bti623}.

@@

{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}

\caption{The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C). Black line represents the model estimation and shaded green the confidence intervals}\label{fig:Figure-5}
\end{figure}

### 2021-08-27 data analysis/sex_analysis.Rmd
cx

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting),
         doyh = yday(harvest)) %>% 
  mutate_if(is_character, as_factor) %>% 
  filter(!is.na(gender)) %>% 
#  filter(!is.na(weight)) %>% 
  mutate(fem = as_factor(if_else(gender == "f", 1, 0))) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) %>% 
  filter(location != "grant") %>% 
  dplyr::select(fem, gender, location, year, month, crop, height, weight, doyh) -> data2
  na.omit(data2) -> data2
```

@@

```{r}
doy_f +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(170, 270), breaks = seq(170, 270, 20)) +
  theme_test() +
  geom_emoji(aes(x = 220, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") + 
    theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> doyf

@@

```{r}
# Probabilities of female with weight
visreg(model_logi, "weight", scale="response", rug=2, gg = TRUE,
       xlab=expression(paste("Weight (g plant "^"-1",")")),
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> weight_f
```
```{r}
weight_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  scale_x_continuous(limits = c(0, 1100), breaks = seq(0, 1100, 150)) +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
    theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> weightf

@@

```{r}
height_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
    theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> heightf

@@

doyf + weightf + heightf +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 30, hjust = 0, vjust = 0))

ggsave("figures/Figure 5.png", width = 9, height = 3)
```

>>>>

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting),
         doyh = yday(harvest)) %>% 
  mutate_if(is_character, as_factor) %>% 
  filter(!is.na(gender)) %>% 
#  filter(!is.na(weight)) %>% 
  mutate(fem = as_factor(if_else(gender == "f", 1, 0))) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) %>% 
  filter(location != "grant") %>% 
  dplyr::select(fem, gender, location, year, month, crop, height, weight, doyh) -> data2
```

@@

```{r}
doy_f +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(170, 270), breaks = seq(170, 270, 20)) +
  theme_test() +
  geom_emoji(aes(x = 220, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") + 
    theme(legend.position = "none"#,
#        axis.text.y = element_markdown(size = 25),
#        axis.title = element_markdown(size = 30),
#        axis.text.x = element_markdown(size = 25),
#        legend.text = element_markdown(size = 25)
        ) -> doyf

@@

```{r}
# Probabilities of female with weight
visreg(model_logi, "weight", scale="response", rug=2, gg = TRUE,
       xlab=expression(paste("Weight (g plant "^"-1",")")),
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> weight_f
```
```{r}
weight_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  scale_x_continuous(limits = c(0, 1100), breaks = seq(0, 1100, 150)) +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
    theme(legend.position = "none"#,
#        axis.text.y = element_markdown(size = 25),
#        axis.title = element_markdown(size = 30),
#        axis.text.x = element_markdown(size = 25),
#        legend.text = element_markdown(size = 25)
        ) -> weightf

@@

```{r}
height_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
    theme(legend.position = "none"#,
#        axis.text.y = element_markdown(size = 25),
#        axis.title = element_markdown(size = 30),
#        axis.text.x = element_markdown(size = 25),
#        legend.text = element_markdown(size = 25)
        ) -> heightf

@@

doyf + weightf + heightf +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 12, hjust = 0, vjust = 0))

ggsave("figures/Figure 5.png", width = 10, height = 4)
```
